---
title: "Cyclistic Report Through Divvy Bikes 2024"
author: "Ferry"
date: "2026-02-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("dplyr")
library("ggplot2")

all_trips <- read_csv("D:/Learn/Data Analysis/Coursera/Case Study/BikeSharing/Data/cleaned_2024/all_trips_processed_2024.csv")

# Add day of week when the trip record is appearing
correct_level <- c(
  "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"
)

all_trips <- all_trips %>% 
  mutate(
    day_of_week = factor(day_of_week, levels = correct_level)
  )
```

## Business Context

#### Situation
Cyclistic is a bike-share company seeking to **maximizing the number of annual memberships**. The marketing team wants to understand **how casual riders and annual members use Cyclistic bikes differently** to support their conversion strategy.

#### Objective
Understand how annual members and casual riders use Cyclistic bike differently based on ride history.

#### Key Questions
1. What is the proportion of rides made by annual members and casual riders in the past year?
2. What patterns differentiate between annual members and casual riders?

<br>

## Ride Distribution
```{r echo=FALSE}
total_proportion <- all_trips %>% 
  count(member_casual) %>% 
  mutate(
    percentage = scales::percent(n / sum(n))
  )
total_proportion %>% 
  ggplot(aes(x = "", y = percentage, fill = member_casual)) + 
  geom_bar(stat = "identity") +
  coord_polar("y") +
  geom_text(aes(label = percentage),
            position = position_stack(vjust = 0.5),
            size = 5,
            fontface = "bold",
            color = "white") +
  scale_fill_manual(values = c("member" = "#1F3A8A", "casual" = "#F97316")) +
  labs(
    title = "Ride Distribution by Rider Type",
    fill = "Rider Type"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold"),
  )

```

Over the past year, 37% of rides were made by casual riders, with 63% of rides by annual members. This indicates that **annual members are dominant in ride usage**. However, casual riders still make up for more than a third by usage, which opens for conversion opportunities.

<br>

## Median Ride Duration
```{r echo=FALSE}
hist(
  all_trips$ride_length, 
  breaks = 50, 
  xlab = "Ride Length", 
  main = "Ride Length Distribution"
  )
```

Ride length data is right-skewed due to the presence of extreme long rides, resulting in a noticeable difference between mean and median values. Therefore, median is used to represent typical ride behavior and provide a fair comparison between casual riders and annual members.

```{r echo=FALSE}
median_distribution <- all_trips %>% 
  group_by(member_casual, day_of_week) %>% 
  summarise(
    Median = median(ride_length, na.rm = TRUE),
    .groups = "drop"
  )

median_distribution %>% 
  ggplot(aes(fill = member_casual, x = day_of_week, y = Median)) +
  scale_fill_manual(values = c("member" = "#1F3A8A", "casual" = "#F97316")) +
  geom_bar(position = "dodge", stat = "identity") +
  labs(
    title = "Median Ride Duration by Day of Week",
    fill = "Rider Type",
    x = "Day of Week",
    y = "Ride Length Median (Seconds)"
  ) +
  theme(
    plot.title = element_text(face = "bold")
  )
```

**casual riders tend to take longer ride duration** on median, indicating a more relaxed usage, whereas **annual members are more efficient**.

<br>

## Weekly Ride Frequency
```{r echo=FALSE}
ride_frequency <- all_trips %>% 
  group_by(member_casual, day_of_week) %>% 
  summarise(
    total_rides = n(),
    .groups = "drop"
  )

ride_frequency %>% 
  ggplot(aes(fill = member_casual, x = day_of_week, y = total_rides)) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(values = c("member" = "#1F3A8A", "casual" = "#F97316")) +
  labs(
    title = "Number of Rides by Day of Week",
    fill = "Rider Type",
    x = "Day of Week",
    y = "Rides"
  ) +
  theme(
    plot.title = element_text(face = "bold")
  )
```

**Annual members demonstrate a consistent weekday** riding pattern, peaking on Wednesday (Mid-week), which **strongly suggests routine commuting behavior**. In contrast, **casual riders show significantly higher rides on weekends**, peaking on Saturday, indicating **leisure-oriented usage**.

